Reducing Power Consumption in Huffman Coding

### 2015-02-23 Switche Paper 1st approache/elsarticle-template-num.tex
exprimernts+ conlusion

The size of biological data base increase with an expanding rate and it doesn't stop for increasing. In the last decade the use of biological data history increase which evolve a high request to data from this high volume biological data base. The exponential growth of these databases become a big problem to all biological data processing methods. Data compression methods is the most used methods to evolve the reduce the cost of treatment of high volume data bases.  The main objective of data compression methods is minimising the number of bits in the data representation. Authors in [] proposed a new approach to genomic data compression based in suffix and common substring and code the difference with Huffman code. In [16], authors propose a general scheme for coding only the difference between the target genome and the referential one, the coding method used is Huffman. Recent work on data compression for biological data have been proposed to deal with transmission of genomic data as an email attachment [][][]. Cost of data transmission is based on two point, firstly the size of the data which has been improved considerably in the recent year with the improvement of Huffman code. Secondly, is the number of switches on the generated codes. The power consumption increase linear with the number of switches.

@@

The crossover operation aims to generate new solutions from the current solution to built the second generation of the population. Furthermore, the mutation provide the diversity on the search space solution by changing the position of  nodes on the same tree. After these two operations, the next step of the genetic algorithm is two to breed the new population. Firstly the new solutions (children) are added to the current population, after that the population is ranked by fitness. The algorithm remove the worst solutions until the initial size of the population is achieved. The algorithm genetic repeat these operators until the stopped criteria is achieved, in this algorithm the stopped criteria is when the objective function (number of switches) stop decreasing.

@@

@@

\begin{tabular}{c  c c  c c  c c}
\hline
 & Huffman Algorithm & P \\\hline
\\\hline
Genome 1& 18.16 & 10.05 & 44.65 \\\hline
Genome 2& 24.66 &  15.63 & 36.61 \\\hline
Genome 3&18.46 &  10.66&  42.25\\\hline
Genome 4&25.18&16.26& 35.42\\\hline
Genome 5& 19.24& 16.08 &16.42 \\\hline
Genome 6& 19.61&14.96&23.71\\\hline
Genome 7& 22.40 &13.21&41.02\\\hline
Genome 8&23.87 & 17.70 &25.84\\\hline
Genome 9&17.66 &10.04&43.14\\\hline
Genome 10& 18.14 &9.89& 45.47 \\\hline
Genome 11&19.63&11.48&41.51 \\\hline
Genome 12& 17.59&11.86&32.57 \\\hline
Genome 13& 17.39&10.93&\\\hline
Genome 14&18.53 &11.64& \\\hline
Genome 15& && \\\hline
Genome 16&&& \\\hline
Genome 17& && \\\hline
Genome 18& && \\\hline

%\bottomrule
\end{tabular}
%\end{center}
\end{table}
\section{Conclusion}

>>>>

The size of biological data base increase with an expanding rate and it doesn't stop for increasing . In the last decade the use of biological data history increase which evolve a high request to data from this high volume biological data base []. The exponential growth of these databases become a big problem to all biological data processing methods. Data compression methods is the most used methods to evolve the reduce the cost of treatment of high volume data bases.  The main objective of data compression methods is minimising the number of bits in the data representation. Authors in [] proposed a new approach to genomic data compression based in suffix and common substring and code the difference with Huffman code. In [16], authors propose a general scheme for coding only the difference between the target genome and the referential one, the coding method used is Huffman. Recent work on data compression for biological data have been proposed to deal with transmission of genomic data as an email attachment [][][]. Cost of data transmission is based on two point, firstly the size of the data which has been improved considerably in the recent year with the improvement of Huffman code. Secondly, is the number of switches on the generated codes. The power consumption increase linear with the number of switches.

@@

The crossover operation aims to generate new solutions from the current solution to built the second generation of the population. Furthermore, the mutation provide the diversity on the search space solution by changing the position of  nodes on the same tree. After these two operations, the next step of the genetic algorithm is two to breed the new population (see fig.2). Firstly the new solutions (children) are added to the current population, after that the population is ranked by fitness. The algorithm remove the worst solutions until the initial size of the population is achieved. The algorithm genetic repeat these operators until the stopped criteria is achieved, in this algorithm the stopped criteria is when the objective function (number of switches) stop decreasing.

@@

\includegraphics[width=350pt,height=170pt]{Images/Drawing3.jpg}
\caption{Population update for genetic algorithm}
\label{Fig2}

@@

\begin{tabular}{c  c c  c c  c c}
\hline
 & Huffman Algorithm & P \\\hline
\\\hline
Genome 1& 18.16 & 10.05 & 44.65 \\\hline
Genome 2& 24.66 &  15.63 & 36.61 \\\hline
Genome 3&18.46 &  10.66&  42.25\\\hline
Genome 4&25.18&16.26& 35.42\\\hline
Genome 5& 19.24& 16.08 &16.42 \\\hline
Genome 6& 19.61&14.96&23.71\\\hline
Genome 7& 22.40 &13.21&41.02\\\hline
Genome 8&23.87 & 17.70 &25.84\\\hline
Genome 9&17.66 &10.04&43.14\\\hline
Genome 10& 18.14 &9.89& 45.47 \\\hline
Genome 11&19.63&11.48&41.51 \\\hline
Genome 12& 17.59&11.86&32.57 \\\hline
Genome 13& 17.39&10.93&37.14\\\hline
Genome 14&18.53 &11.64&37.18 \\\hline
Genome 15&78.19 &50.60&35.28 \\\hline
Genome 16& & & \\\hline
Genome 17& & & \\\hline
Genome 18& & & \\\hline
%\bottomrule
\end{tabular}
%\end{center}
\end{table}
\section{Conclusion}
In this paper we described how the genetic algorithm can be used to improve the power consumption in biological data transmission. The advantage of the proposed approach is that it minimise considerably the number of switches with affect directly the power consumption for data transmission. In this paper, we identified genetic
algorithm as a potential method to optimise the total number of inter- and intra-switches. The Genetic algorithm start by generating a set of feasible solutions (trees), each one contain a set of codewords. After that, the approach uses the natural genetic operators to improve the quality if the solutions by generating new solution. the evolution of the population is controlled by the number of switches both inter- and intra-switches. 
The proposed approach is applied to biological data sets with different sizes from small genome to a human chromosome. The approach has been compared with the classic Huffman code. The results show that
the proposed approach improves the power consumption rate considerably. The experiment shows that the proposed approach improves Huffman code in the best case with $best$,  $average$ in the average case, and $worst$ case. For the future we hope to improve the allocation of codewords to different frequencies in orderer to improve the total number of switches.
